---
title: "HW: Week 4"
author: "36-350 -- Statistical Computing"
date: "Week 4 -- Spring 2022"
output:
  pdf_document:
    toc: no
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes
---

Name: christina choi

Andrew ID: gawonc

You must submit **your own** HW as a PDF file on Gradescope.

```{r wrap-hook,echo=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

---

## Question 1
*(40 points)*

You are given the following matrix:
```{r linewidth=80}
set.seed(505)
mat = matrix(rnorm(900),30,30)

mat[sample(30,1),sample(30,1)] = NA


```


Compute the standard deviation for each row, using `apply()` and your own on-the-fly function, i.e., a function that is defined *within* the argument list being passed to `apply()`. **Do not use the function sd()!** Realize that since there is a missing value within the matrix, you need to define your function so as to only take into account the non-missing data in each row. If your vector of standard deviations has an `NA` in it, then your function isn't quite working yet.

```{r linewidth=80}

apply(mat, MARGIN = 1, FUN= function(x) { 
     x=x[!is.na(x)]
     sumof = sum((x - mean(x))^2)
     sqrt(sumof/(length(x)))
  
  })

```
---

Below we read in the data on the political economy of strikes.
```{r linewidth=80}
strikes.df = read.csv("http://www.stat.cmu.edu/~mfarag/350/strikes.csv")
```

---

## Question 2
*(40 points)*

Using `split()` and `sapply()`, compute the average unemployment rate, inflation rates, and strike volume for each year represented in the `strikes.df` data frame. The output should be a matrix of dimension 3 $\times$ 35. (You need not display the matrix contents...just capture the output from `sapply()` and pass that output to `dim()`.) Provide appropriate row names (see `rownames()` to your output matrix. Display the columns for 1962, 1972, and 1982. (This can be done in one line as opposed to three.)
```{r linewidth=80}

three.calc = function(strikes.df) {
  return(c(mean(strikes.df[[3]]), mean(strikes.df[[4]]), mean(strikes.df[[5]] )))
}

s = split(strikes.df, strikes.df$year)
s
date.split =sapply(s, FUN=three.calc )

dim(date.split)

rownames(date.split) = c("strike volume",  "unemployment", "inflation rates" )

date.split[, c("1962", "1972", "1982")]



```

## Question 3
*(40 points)*

Utilize piping and `group_by()`, etc., to compute the average unemployment rate for each country, and display that average for only those countries with the maximum and minimum averages. To be clear: your output should only show average unemployment for Ireland and Switzerland, and nothing else. (Hint: remember `slice()`, a less-often-used `dplyr` function.) Hint: arrange your output in order of descending average unemployment, then note that `n()` applied as an argument to the right function will return the last row.

```{r linewidth=80}
library(magrittr)
library(tidyverse)

new.strikes.df = strikes.df %>% group_by(country) %>%
  select(.,unemployment) %>% summarize(mean.unemployment = mean(unemployment)) %>%
  arrange(mean.unemployment) 

rbind(slice_head(new.strikes.df), slice_tail(new.strikes.df))




```

